---
title: "Exam III -- Takehome Exam"
output: html_document
---

## Instructions

Due: Start of class, Thursday 4 April 2019

Write code as needed below and fill in your written answers in full
sentences wherever you see the prompt "Answer:". I have sent you a folder on
Box to upload your solutions to. Please upload the Rmd file and a knitted html
file with your responses. If you have trouble with the upload, come to class
on Thursday with the completed assignment and I will assist with the upload
process.

Note: If you believe that there is something unspecified in the question, any
way of following the instructions will be marked correctly.

## Packages

Read in the following R packages that we will need for the exam:

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(smodels)
library(tmodels)
library(readr)

theme_set(theme_minimal())
```

## Another Weather Dataset

We are going to look at a NASA weather dataset. This particular
collection contains information about Atlantic storms. Read it in with the
following (it is also doing some pre-processing; you do not need to understand
it, just run the entire chunk of code):

```{r}
storms <- read_csv("https://statsmaths.github.io/stat_data/storms.csv", guess_max = 10000)
storms <- mutate(storms, name = paste(year, name, sep="-"))
storms <- mutate(storms, date = as.Date(sprintf("%04d-%02d-%02d", year, month, day)))
```

It will also be useful to have a dataset giving the borders of countries:

```{r}
borders <- read_csv("https://statsmaths.github.io/stat_data/nasa_borders.csv")
```

1. Draw a bar plot showing the number of storm observations in each year:

```{r}

```

What year seems to have had the most activity?

**Answer**:

2. Filter the dataset to only include storms that have a status of a hurricane. Construct
a histogram of the wind speeds of these storms.

```{r}

```

3. Filter the storms data to create a dataset only containing measurements from the
2013 storm Barry ("2013-Barry"). Draw a plot with longitude on the x-axis, latitude
on the y-axis. Include two geometries: a path geometry (geom_path) and a point
geometry (geom_point). For the points, color by the windspeed.

```{r}

```

4. Create a plot with the `borders` dataset. Put longitude on the x-axis and latitude
on the y-axis. Include a path geometry and set the `group` aesthetic to the variable
`group` in the dataset. You should see a map of Central America and the lower half of
the United States.

```{r}

```

5. Combine the plots from questions 3 and 4 (not literally; make a new
plot that contains both the map and the path of hurrican Barry). Make sure that
the hurricane data is on top of the map.

```{r}

```

What country seems to be have been most effected by hurricane Barry in 2013?

**Answer**:

6.Take the plot from the previous question, but replace the path geometry with the
map with a polygon geometry (geom_polygon). Set the `alpha` aesthetic to 0.2 to make
the map opaque.

```{r}

```

7. Construct a dataset of all storms that occured in 1995. Repeat the plot from the
prior question, but this time using all of the storms from 1995. To keep the plot from
becoming too clutered add two additional layers: (i) a facet wrap by the name of the storm
and (ii) the layer theme_void().

```{r}

```

8. Create a new dataset consisting of only those measurements in 1995 that had a status
equal to a hurricane. Starting with the plot from the previous question, add a layer of
red points showing where each storm attained hurricane strength.

```{r}

```

9. Construct a plot with month on the x-axis and wind (the speed of the storm) on the y-axis.
Add a `geom_mean` layer to the plot to see the average wind speed of storms in each month.

```{r}

```

What month or months seems to have the storms with the highest wind speeds?

**Answer**:

10. Draw a plot with month on the x-axis and wind speed on the y-axis. Color the
points based on the storm status.

```{r}

```

What wind speeds seem to be the cutoff for (i) a tropical storm and (ii) a
hurricane?

**Answer**:



