---
title: "Exam IV -- Takehome Exam"
output: html_document
---

## Instructions

Due: Start of class, Thursday 25 April 2019

Write code as needed below and fill in your written answers in full
sentences wherever you see the prompt "Answer:". I have sent you a folder on
Box to upload your solutions to. Please upload the Rmd file and a knitted html
file with your responses. If you have trouble with the upload, come to class
on Thursday with the completed assignment and I will assist with the upload
process.

Note: If you believe that there is something unspecified in the question, any
way of following the instructions will be marked correctly.

**Remember, this exam can be skipped if you are happy with your first three
exam grades.**

## Packages

Read in the following R packages that we will need for the exam:

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(smodels)
library(tmodels)
library(readr)

theme_set(theme_minimal())
```

## Bike Rentals in DC

Bike sharing systems are variants of traditional bicycle rentals, where
the process of renting and returning is heavily automated; typically, bikes
can be rented at one location and returned at another without ever having
to deal with a human being. There are currently several hundred bike sharing
programs in many different cities, and a great deal of interest in their
potential.

For this data analysis, we have a set of data from a bike sharing program
derived from a two-year usage log of a Washington, D.C. bike-sharing system
called Captial Bike Sharing (CBS). You can read the data in with the following

```{r}
bikes <- read_csv("https://statsmaths.github.io/stat_data/bikes.csv")
bikes <- mutate(bikes, date = as.Date("2011-01-01") + seq_len(nrow(bikes)) - 1L)
```

The following fields are available for you to work with:

-  season       - season of the year, 1 to 4 (Winter, Spring, Summer, Fall)
-  year         - code for the year: 0=inital year of program, 1=second year
-  workingday   - 1 for working days, 0 for weekends and holidays
-  weather      - weather, coded as follows:
    - 1 = Clear to partly cloudy
    - 2= Mist but no heavier precipitation
    - 3= Light rain or snow, possibly with thunder
    - 4= Heavy rain or snow
-  temp         - temperature in Celsius
-  humidity     - humidity percentage
-  windspeed    - windspeed in kilometers per hour
-  registered   - number of bike rentals that day by registered users
-  count        - count of total bike rentals that day, including both casual
                  and registered users
-  date         - a data classed object that contains the day of the dataset;
                  created in the code block above

1. Using pipes, draw a scatterplot with the number of bikes registered
on the y-axis and the tempurature on the x-axis for just the initial year
of the program.

```{r}

```

2. Approximately, what seems to be the ideal tempurature for biking in DC?

**Answer**:

3. Create a new variable called `temp_f` that gives the tempurature in fahrenheit
(you need look up the formula for converting from celcius). Then, redo the plot from
question 1 with `temp_f` on the x-axis. Do all of this in a single sequence of commands
with a pipe staring with the full dataset bike.

```{r}

```

4. Using a pipe along with the `arrange` and select function, sort the
dataset in descending order by the total number of bikes rented and
select just the variables count and data.

```{r}

```

On what date were the highest number of bikes rented?

**Answer**:

5. Using a single pipe function, create a new variable called
`year_season` that pastes together year and season (paste(year, season, sep="-")).
Then, summarize the average, minimum, and maximum number of bikes
and construct a plot with `year_season` on the x-axis and number
of bikes on the y-axis. Show three lines (geom_line): one for the
minimum, one for the maximum, and one for the average value.
Somehow differentiate the max/min from the average value (you can use
linetype, color, or line size to do this; completely up to you).

Note: In order for this to work, the first line of your plot needs to
be this (without the quotes): `ggplot(aes(year_season, count_avg, group=1)) +`

```{r}

```

6. Given a date object in R, you can extract a string giving the year and month using
the format command. For example, take this code:

```{r}
format(as.Date("2019-04-15"), "%Y-%m")
```

Create a new dataset called `bikes_month` by constructing a variable year_month that
extracts the year and month from the date variable in the dataset, grouping by the
year_month variable, and summarizing by adding a variable called `avg_bikes` to get
the average number of bikes rented in each time period. Do all of this with a single
pipe command.

```{r}

```

7. Using the left_join function, join the information from the `bikes_month` dataset.
Note: you will have to reconstruct the year_month variable in the bikes dataset. Pipe
this joined dataset into a plot that shows the average number of bikes rented in each
month (y-axis) as a function of the date variable using a line and the actual counts
of the bikes as points. Change the color of line to something other than black and make
its size equal to 3. Set the opacity of the points (alpha) to 0.3 and make sure that the
points are behind the line.

```{r}

```

8. The following code snippet illustrates the usage of the `if_else` function. It creates
a new variable based on whether a logical statement is ture or false; the code apply this
function to the numbers 1 through 10

```{r}
if_else(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10) > 5, "Large number", "Small number")
```

Modify the plot from the previous question to make the color of the points change based on
whether they are below the line of the average monthly count or above the line. Do this using
a single pipe command and the if_else function.

```{r}

```

9 This last question requires you to apply two new functions: lag
and lead. They are used within a mutate function in order to allow the construction
of new variables based on the values of the previous and subsequent data rows. They
are particularly useful for datasets such as the bikes data that have a sequence of
observations over time. You can look here for more information about these functions:

- https://dplyr.tidyverse.org/reference/lead-lag.html

Using a set of pipes, construct a variable named count_prev and count_next to grab
the number of bikes rented in the previous and next days, respectively. Then create
a variable avg_before_after that is equal to the average of these two counts. That
is, (count_prev + count_next) / 2. Finally, pipe this into a plot that shows the
relationship between the average number of bikes on the subsequent days (x-axis) and
the number of bikes rented on the day itself (y-axis). Add a regression line to the
plot. Note: you may get a warning that two rows were removed from the dataset. This
is normal and not somethign to worry about.

```{r}

```

10. There is nothing for you to do for this question. Your grade for question 10 depends
on how well your code is formated throughout the exam. Make sure all pipes start with just
the dataset and that your plots and variables follow the conventions from the course.




