---
title: "Final Project: Exploring A Week in the Life"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this project, you need to describe the dataset that you have collected
and then write a short (approx. 500 words with 3-5 graphics) exploratory
analysis of the dataset. In order to get an 'A' on the assignment, you should
integrate some of the advanced graphics and data manipulation techniques that
we have been learning. There is no need, however, to exhaustively use every
technique; only apply what makes sense for your dataset.

To submit, upload your three datasets, this script, and a knitted version of
the HTML file. The project is due before class on Thursday, 25 April.

## Data dictionary

Table of hourly activities :

- **name**:
- **year**:
- **day**:
- **month**:
- **hour**:
- **location**:
- **activity**:
- **new_var**:

Table of locations:

- **location**:
- **lat**:
- **lon**:
- **new_var**:

Table of hourly activities:

- **activity**:
- **score**:
- **new_var**

## Read in the datasets

The code below reads in your datasets. It also attaches a date-like object
to the hourly table that may be useful in your plots:

```{r}
hourly <- read_csv2("hourly.csv")
activities <- read_csv("activities.csv")
places <- read_csv("places.csv")

hourly <- mutate(hourly,
                 time = as.POSIXct(sprintf("%04d-%02d-%02d %02d:00", year, month, day, hour)))
```


## Analysis


