---
title: "Lab 08"
output: html_document
---

Start by reading in all of the R packages that we will need for today:

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(tmodels)
library(readr)
library(readxl)
```

## Class data collection

We are going to build a dataset as a class. Download the datafile and save
it as "class08.csv". Read in the dataset and call it simply `class` using
the `read_csv` function:

```{r}

```

Run all three correlation tests on the data:

```{r}

```

Are there any large differences between the p-values in the tests? What do you
conclude from the analysis about the relationship between the variables?

## Graduate pay information by major

Now, read in the pay by major dataset that we used last time:

```{r, message=FALSE}
pay <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/grad_info.csv")
```

Run Pearson's product-moment correlation test on this dataset with `median_pay`
as the response and `unemployment` as the independent variable. Note the correlation
point estimate and the p-value:

```{r}

```

Re-run the analysis with `median_pay` as the independent variable and `unemployment`
as the response.

```{r}

```

The results should be exactly the same due to symmetry in the test.

## Graduate pay by major in three categories

The following code creates a new variable in the dataset. Run it and look at
the dataset to see what it does

```{r}
pay$mcategory <- "non-science"
pay$mcategory[pay$sciences == "yes"] <- "other science"
pay$mcategory[grep("ENGINEERING", pay$major)] <- "engineering"
```

The variable `mcategory` now has three different values: non-science, other
science, and engineering. Use the one-way ANOVA function to test a relationship
between median pay and the variable `mcategory`:

```{r}

```

Now, run the same test with Kruskal-Wallis rank sum test:

```{r}

```

Finally, make a copy of the dataset and introduce a bad data point as we
did last time:

```{r}
pay_bad <- pay
pay_bad$median_pay[173] <- 22000000
```

Verify that the one-way ANOVA function is sensitive to this outlier but the
Kruskal-Wallis test is not:

```{r}

```










