---
title: "Lab 15: Annotating Data Visualizations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
library(smodels)
theme_set(theme_minimal())
```


## Today

Today you'll be working with three different datasets. For each, you'll
need to produce a single plot with labels and annotations. You should make
use of several manual annotations in an effort to tell a visual story about
the dataset. This will be good practice for the next exam.

### Murder rates

The first dataset gives murder rates across different cities in the US:

```{r}
murder <- read_csv("https://statsmaths.github.io/stat_data/murder_by_city.csv")
```

Create a plot that visualizes this dataset. The specifics are up to you,
but it must include at least 2 manual annotations, one aesthetic, and
labels axes.

```{r}

```

### DJIA

The second dataset for data are daily close values of the Dow Jones Industrial
Average:

```{r}
djia <- read_csv("https://statsmaths.github.io/stat_data/djia_daily.csv")
```

Create a line plot with time on the x-axis and the close price on the y-axis.
Then add axis labels and three manual annotations.

Note that you will have to use the format `x = as.Date("2017-01-06")` to add
annotations to an axis that uses time as a component.

```{r}

```

### Presidential approval

The final dataset investigates Donald Trump's approval rating:

```{r}
trump <- read_csv("https://statsmaths.github.io/stat_data/trump_approval.csv")
```

Create a plot the simultaneously shows the approval and disapproval ratings
(you'll need two `geom_line` layers). Also label the axes and include three
manual annotations.

```{r}

```

Note that you will have to use the format `x = as.Date("2017-01-06")` to add
annotations to an axis that uses time as a component.


