---
title: "Lab 10: Logistic Regression"
output: html_document
---

I needed to make some further changes to **tmodels**. Reinstall with the
following code:

```{r}
devtools::install_github("statsmaths/tmodels")
```

Then, read in all of the R packages that we will need for today:

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(tmodels)
library(readr)
```

## NBA Dataset

Today we are going to look at a dataset from the NBA. Specifically, trying to
see what factors influence whether a shot is missed or made. Run the following
code to load the dataset. Do not worry about what the extra lines of code are
doing; I am just cleaning the dataset to make it easier to model. You will learn
how to do that in the second half of the course.

```{r, message=FALSE}
nba <- read_csv("https://statsmaths.github.io/ml_data/nba_shots.csv")
nba <- filter(nba, !is.na(fgm))
nba <- mutate(nba, shot = if_else(fgm == 0, "Missed", "Made"), period = sprintf("p%d", period), pts_type = if_else(pts_type == 2, "two", "three"))
nba <- select(nba, shot, pts_type, period, shot_clock, dribbles, touch_time, shot_dist, close_def_dist, shooter_height)
```

You can see the data with this command:

```{r}
nba
```

1. Let's first see how well players are able to make shots based on whether it is
two or three point attempt. Build a contingency table of the variables `shot` (response)
and `pts_type`:

```{r}

```

2. I want you to use of the contingency table tests that we had for the first exam on
the shots as a function of points type contingency table. Pick the test that seems
the most appropriate and run it below:

```{r}

```

Is the test significant? What is the point estimate and what does it mean?
Does the sign of the point estimate (negative or positive) make sense?

3. Re-do the analysis with logistic regression.

```{r}

```

You should see that the logistic regression flips the classes due to the
internal logic of the **tmodels** package. Notice that the point estimate
is different from the T-test and understand why we would not expect these
numbers to be the same.

4. Instead of only using shot type, let's use instead the variable `shot_dist`
and see how it effects the shot type. You could use these two variables with
a two-sample T-test like this:

```{r}
tmod_t_test(shot_dist ~ shot, data = nba)
```

Explain why this test is not appropriate for our application.

5. Use logistic regression to instead investigate how the likelihood of a shot
being made changes with shot length:

```{r}

```

Does the sign of the point estimate (negative or positive) make sense given what
we are predicting here?

6. Run a logistic regression that predicts whether a shot was made based on the
distance to the closest defender.

```{r}

```

You should see that the p-value is less than 0.05 but not very small (larger
than 0.01, for example). Does the sign of the point estimate make sense to you?

7. Re-run the logistic regression in question 6 but include shot distance as a
nuisance variable:

```{r}

```

How does this nuisance variable change the p-value and the point estimate?
Can you explain / summarize what might be going on here?

8. A friend claims that the shot clock (time left for a player to
attempt a shot) is on average half-way finished when the average shot
is made. The NBA shot clock lasts 24 seconds. Run a one-sample T-test to
test this hypothesis:

```{r}

```

Does the p-value suggest that your friend is likely incorrect? Looking at
the point estimate, however, would it be fair to say that your friend is
completely mistaken?






