---
title: "Lab 05"
output: html_document
---

In this file, I've tried to explain some of the best practices for how to
approach the problems in the lab file. To start, I recommend loading all of
the libraries you will need right here at the top:

```{r}
library(readr)
library(readxl)
library(tmodels)
```

**Question 1**

This is an observational study because the researchers were not able to manually
assign players to being in the "tall" or "short" group. Therefore, we cannot
prove a causal relationship by this study.

**Question 2**

The author's mostly interpret the results correctly by hedging their conclusions
with phrases such as "predictive value", "may impair", and "potentially increasing".

**Question 3**

See the Excel file lab05-table3.xlsx, which is posted online.

**Question 4**

I saved the dataset as an Excel file (you can find a link on the class website),
and placed it in the same location as the RMarkdown file. Here is the code to read
in the dataset and show the first few rows in RStudio:

```{r}
hlas <- read_xlsx("lab05-table3.xlsx")
hlas
```

While not required, here is the contingency table to check against the paper:

```{r}
tmod_contingency(las ~ height, data=hlas)
```

And, here is the statistical hypothesis test:

```{r}
tmod_fisher_test(las ~ height, data=hlas)
```

Amazingly, this is not the p-value reported in the paper (which is given as
0.04)!

**Question 5**

The other tests are given by (the warning about the Chi-squared test is not something
to worry about for now):

```{r}
tmod_chi_squared_test(las ~ height, data=hlas)
tmod_z_test_prop(las ~ height, data=hlas)
```

These p-values are quite a bit different than the other tests. While they tend to
be close in general, they diverge when some of the counts in the contingency tables
are very small.

**Question 6**

Just given the assumptions, the best model is the chi-squared because neither the row
nor the column sums are fixed ahead of time.

**Question 7**

Now, read in the second set of data:

```{r}
plas <- read_xlsx("lab05-table4.xlsx")
plas
```

And again, here is the contingency table for us to verify with the paper:

```{r}
tmod_contingency(las ~ previous_sprain, data=plas)
```

Finally, here are the three tests:

```{r}
tmod_z_test_prop(las ~ previous_sprain, data=plas)
tmod_chi_squared_test(las ~ previous_sprain, data=plas)
tmod_fisher_test(las ~ previous_sprain, data=plas)
```

Notice once again that the test does not match the paper! This is not
good, but also not our fault.

