---
title: "Lab 10: Logistic Regression"
output: html_document
---

I needed to make some further changes to **tmodels**. Reinstall with the
following code:

```{r}
# devtools::install_github("statsmaths/tmodels")
```

Then, read in all of the R packages that we will need for today:

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(tmodels)
library(readr)
```

## NBA Dataset

Today we are going to look at a dataset from the NBA. Specifically, trying to
see what factors influence whether a shot is missed or made. Run the following
code to load the dataset. Do not worry about what the extra lines of code are
doing; I am just cleaning the dataset to make it easier to model. You will learn
how to do that in the second half of the course.

```{r, message=FALSE}
nba <- read_csv("https://statsmaths.github.io/ml_data/nba_shots.csv")
nba <- filter(nba, !is.na(fgm))
nba <- mutate(nba, shot = if_else(fgm == 0, "Missed", "Made"), period = sprintf("p%d", period), pts_type = if_else(pts_type == 2, "two", "three"))
nba <- select(nba, shot, pts_type, period, shot_clock, dribbles, touch_time, shot_dist, close_def_dist, shooter_height)
```

You can see the data with this command:

```{r}
nba
```

1. Let's first see how well players are able to make shots based on whether it is
two or three point attempt. Build a contingency table of the variables `shot` (response)
and `pts_type`:

```{r}
tmod_contingency(shot ~ pts_type, data = nba)
```

2. I want you to use of the contingency table tests that we had for the first exam on
the shots as a function of points type contingency table. Pick the test that seems
the most appropriate and run it below:

```{r}
tmod_z_test_prop(shot ~ pts_type, data = nba)
```

Is the test significant? What is the point estimate and what does it mean?
Does the sign of the point estimate (negative or positive) make sense?

**Yes, it is significant (pvalue: < 2.2e-16). The point estimate gives
the increased probability of making a two-point shot compared to a three-point
shot. It is positive, which makes sense because two-point shots are easier
to make.**

3. Re-do the analysis with logistic regression.

```{r}
tmod_logistic_regression(shot ~ pts_type, data = nba)
```

You should see that the logistic regression flips the classes due to the
internal logic of the **tmodels** package. Notice that the point estimate
is different from the T-test and understand why we would not expect these
numbers to be the same.

**It is different because the point estimate here indicates the increase
(decrease here) in the log-odds rather than the probability itself.**

4. Instead of only using shot type, let's use instead the variable `shot_dist`
and see how it effects the shot type. You could use these two variables with
a two-sample T-test like this:

```{r}
tmod_t_test(shot_dist ~ shot, data = nba)
```

Explain why this test is not appropriate for our application.

**It is not appropriate because the test assumes that the shot success causes
the shot distance, rather than the other way around.**

5. Use logistic regression to instead investigate how the likelihood of a shot
being made changes with shot length:

```{r}
tmod_logistic_regression(shot ~ shot_dist, data = nba)
```

Does the sign of the point estimate (negative or positive) make sense given what
we are predicting here?

**The point estimate is positive, indicating that the shot is more likely to be
misseed as we move away from the basket. This seems reasonable because longer
baskets are more difficult to make.**

6. Run a logistic regression that predicts whether a shot was made based on the
distance to the closest defender.

```{r}
tmod_logistic_regression(shot ~ close_def_dist, data = nba)
```

You should see that the p-value is less than 0.05 but not very small (larger
than 0.01, for example). Does the sign of the point estimate make sense to you?

**The point estimate is positive, meaning the the probability of shot=Missed
increases as the distance to the defender increases. This seems a bit counterintuitive
at first; you would think that a closer defender would make it harder to make a 
shot, not easier. Look to the next two questions for an explanation.**

7. Re-run the logistic regression in question 6 but include shot distance as a
nusiance variable:

```{r}
tmod_logistic_regression(shot ~ close_def_dist + shot_dist, data = nba)
```

How does this nusiance variable change the p-value and the point estimate?
Can you explain / summarize what might be going on here?

**The p-value is now very small and the sign of the point estimate has flipped.
After accounting for shot distance, the distance of the defender now works they
way we would expect: a closer defender makes it harder to make a basket. 
Putting this together, the univariate analysis was confounded by the shot length;
shots that were far from defenders were also very long shots far from the basekt.
That is why they looked hard even though there were no nearby defenders.**

8. A friend claims that the shot clock (time left for a player to
attempt a shot) is on average half-way finished when the average shot
is made. The NBA shot clock lasts 24 seconds. Run a one-sample T-test to
test this hypothesis:

```{r}
tmod_t_test_one(shot_clock ~ 1, data = nba, h0 = 12)
```

Does the p-value suggest that your friend is likely incorrect? Looking at
the point estimate, however, would it be fair to say that your friend is
completely mistaken?

**The p-value is significant (p 8.1e-08), however the confidence interval
goes from 12.1s to 12.3 seconds and therefore the value of 12s is not very
far away in practical terms. This was we refer to a statistically significant
result that is not practically significant. Generally we need both to hold
to say that we have an interesting effect.**




