---
title: "Class 21: Relational Data"
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/class21/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

```{r, message = FALSE, include = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(kableExtra)
library(smodels)

theme_set(theme_minimal())
msleep <- na.omit(msleep)
```

## Airline data

Today we will once again look at the NYC flights dataset:

```{r}
flights <- read_csv("https://statsmaths.github.io/stat_data/flights.csv")
```

```{r}
flights <- select(flights, year:day, hour, tailnum, carrier)
```

```{r}
flights
```

This time we will also look at another table that describes each of the
airlines.

```{r}
airlines <- read_csv("https://statsmaths.github.io/stat_data/f_airlines.csv")
airlines
```

These two tables are connected by way of a common shared
variable, known in database theory as *keys*. A key may be
a single variable or a collection of variables (known as
a composite key).

In order to combine them, by matching up along keys, we can make use of the
function `left_join`. Here we see that it returns a new table with the same
number of rows as the `flights_sml` dataset but now with the full name of the
carrier:

```{r, include=FALSE}
flights <- select(flights, year:day, hour, tailnum, carrier)
```

```{r}
left_join(flights, airlines, by = "carrier")
```

And the resulting dataset combines all of the variables by the common key.



