---
title: "Class 19: Collecting Relational Data"
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/class19/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

```{r, message = FALSE, include = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(kableExtra)
library(smodels)

theme_set(theme_minimal())
```

In order to let you get started on the final project, I have reordered the
typical order of my notes on data manipulation. Today we are going to see 
how to collect and manipulate data stored in multiple tables.

## Tidy data

## Three principles

We have already discussed at length how to store tablular datasets and 
some good practices for naming variables. As datasets become more complex,
there are some additional concerns that arise. All of the principles of
constructing a dataset (equivalently, a database) could easily fill a
whole course. Here are three principles that get us on the right track:

- determine the objects of study; each of these gets its own table,
and each example gets its own row; **movies**, **actors**,
**actor-movie links**
- information about an object of study should be put on the relevant
table; the information should remain internally consistent regardless of
changes made within a data table
- each column should be indivisible and the variable type clear;
for example, budget should not include the dollar sign, if needed
create a new column; name columns with no spaces or special characters

If you would like to know about specifics of these principals I highly
recommend the Hadley Wickhams "Tidy Data" paper:

- ["Tidy Data" (2014)](http://vita.had.co.nz/papers/tidy-data.html)

The Tidy Data chapter of R for Data Science covers several useful
functions the show to convert a dataset that has already been collected
into a tidy format:

- [Tidy Data R4DS](http://r4ds.had.co.nz/tidy-data.html)

To better explain these principles, let's look at several tables from the
airline dataset.

## Airline data

Today we will once again look at the NYC flights dataset:

```{r, message=FALSE}
flights <- read_csv("https://statsmaths.github.io/stat_data/flights.csv")
flights
```

This time we will also look at another table that describes each of the
airlines.

```{r, message=FALSE}
airlines <- read_csv("https://statsmaths.github.io/stat_data/f_airlines.csv")
airlines
```

Another describing the planes:

```{r, message=FALSE}
planes <- read_csv("https://statsmaths.github.io/stat_data/f_planes.csv")
planes
```

The airports:

```{r, message=FALSE}
airports <- read_csv("https://statsmaths.github.io/stat_data/f_airports.csv")
airports
```

And the weather:

```{r, message=FALSE}
weather <- read_csv("https://statsmaths.github.io/stat_data/f_weather.csv")
weather
```

Each of these tables captures a specific unit of observation. Information
about a unit of observation is only included in a specific table. Note that,
for example, including the origin airport in the flights table is not an
error: this **is** information about the flight. However, if we put in
information about the timezone of the flight, this would break the tidy
data principles.

## Two table verbs

Having data places across different tables is useful, but often we need to
put the data back together before using it. Notice that the various tables
are connected by way of common shared variables, known in database theory
as *keys*. A key may be a single variable or a collection of variables
(a composite key).

In order to combine them, by matching up along keys, we can make use of the
function `left_join`. Here we see that it returns a new table with the same
number of rows as the `flights` dataset but now with the full name of the
carrier. I'll make use of a pipe to reduce the number of variables to make it
easier to see:

```{r}
left_join(flights, airlines, by = "carrier") %>%
  select(year, month, day, origin, dest, carrier, name)
```

The resulting dataset combines all of the variables by the common key.

### Composite keys

Recall that a key may consist of multiple variables. To
join on many variables at once, provide a vector of all the names
to the option `by`:

```{r}
flights_sml <- flights %>%
  select(year:day, hour, origin, dest)

inner_join(flights_sml, weather,
          by = c("year", "month", "day", "hour", "origin"))
```

### Common variables

Sometimes there may be a variable name present in two datasets that
we want to merge together but that has a different meaning in each
dataset. For example, `year` is the year of the flight in the `flights`
dataset but the year of creating in the `airplanes` dataset. If
we join these two, we see that a suffix is added to each variable:

```{r}
flights_sml <- flights %>%
  select(year:day, tailnum)
inner_join(flights_sml, planes,
          by = c("tailnum"))
```

This behavior is fine, but it is better to manually describe
what the suffix should be:

```{r}
inner_join(flights_sml, planes,
          by = c("tailnum"),
          suffix = c("", "_plane"))
```

### Unspecified key

If we do not specify the key to join with in the `by` option, **dplyr**
will assume that we want to join on all common keys. A warning will be
produced warning which variables were choosen. This can be useful in a
pinch, but generally is a bad idea to rely on.

### Practice (or not)

In the interest of time, we will practice the `left_join` function on the
next lab. We will spend the end of class today discussing the final project.
