---
title: "Class 19: Summarizing Data"
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/class19/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

```{r, message = FALSE, include = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(kableExtra)
library(smodels)

theme_set(theme_minimal())
```

### Learning Objectives

Understand the concepts behind summarizing datasets by a grouping variable
and be able to apply this to a new dataset using the syntax in R.

### NYC Flights Data

Once again we are going to work today with the NYC Flights dataset.

```{r}
flights <- read_csv("https://statsmaths.github.io/stat_data/flights.csv")
flights
```

Take note of the unit of observation here: each row is a flight.

### Changing the Unit of Observation

Often, it is useful to change the unit of observation within a dataset.
The most common way of doing this is to group the dataset by a combination
of variables and aggregate the numeric variables by taking sums, means, or
some other summary statistics. Some common examples include:

- aggregating individual shot attempts in soccer to summary statistics about each player
- aggregating census tract data to a county or state level
- aggregating information about individual patients to summaries about demographic groups

We have seen a few simple ways of doing this already within a plot (such as
counting occurances in a group with `geom_bar`). Today we will see how to do this
with the `group_summarize` command.

### Summarizing data

The group summarize command comes from the **smodels** package. Applying it to a
dataset with no additional options yields a new dataset with just a single line.
Variables in the new dataset summarize the numeric variables in the raw data.

```{r}
flight_line <- group_summarize(flights)
flight_line
```

Specifically, we see the following summaries for each numeric variable (the new names add a suffix
to the original variable name):

- `mean`: the average of all the observations
- `median`: if we ordered all observations from smallest to largest, the middle value
- `sd`: the standard deviation, a measurment of how much the number varies across observations (more on this after the break)
- `sum`: the sum of all the observations

There is also a variable just called `n` at the end, giving the total number of observations in
the entire dataset.

### Group Summarize

The magic of the `group_summarize` command comes from specifying other variables in function call.
If we specify a grouping variable, here I'll use `month`, the summarizing will be done *within*
each month. So, here, the new dataset has 12 rows with each row summarizing a given month:

```{r}
flight_month <- group_summarize(flights, month)
print(flight_month, n = 12)
```

This dataset can then be used in further visualizations. Such as:

```{r}
ggplot(flight_month, aes(month, dep_delay_mean)) +
  geom_line(color = "salmon", size = 1) +
  geom_line(aes(y = arr_delay_mean), color = "olivedrab", size = 1) +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  annotate("text", x = 6.5, y = 23, label = "Delays Peak in July and August...") +
  annotate("text", x = 9, y = -5, label = "... And dip in September") +
  annotate("text", x = 3, y = 15, color = "salmon", label = "Departure") +
  annotate("text", x = 3, y = 4, color = "olivedrab", label = "Arrival") +
  xlab("Month of the Year") +
  ylab("Average Delay in Minutes")
```

Notice that it would be impossible to produce this graphic without the summarize command.

### Summarize By Multiple Variables

By supplying multiple variables to the `group_summarize` command, we can group simultaneously by both.
Here we have a unique row for each combination of carrier and departure airport:

```{r}
flight_carrier <- group_summarize(flights, carrier, origin)
```

Which allows us to make plots like this:

```{r}
ggplot(flight_carrier, aes(carrier, arr_delay_mean)) +
  geom_point(aes(color = origin)) +
  xlab("Carrier Code") +
  ylab("Average Arrival Delay (minutes)")
```

As you can imagine, summarizing data can quickly allow for very complex
graphics and analyses.

