---
title: "Class 22: Statistical Arguments"
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/class22/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

```{r, message = FALSE, include = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(kableExtra)
library(smodels)

theme_set(theme_minimal())
msleep <- na.omit(msleep)
```

## Thesis Statement

You have certainly heard of a thesis statement, but may struggle
to formally define what it is. A thesis statement:

> offers a concise summary of the main point or claim of the essay,
> research paper

A hypothesis, which is not the same concept, is:

> a supposition or proposed explanation made on the basis of
> limited evidence as a starting point for further investigation

That is, a hypothesis is a fixed explanation of something. A
thesis is a particular rhetorical element that exists within a
document or speech and is intended for a particular audience.
If you are designing an experiment or collecting data, you will
have a hypothesis. When you are writing a paper or giving a
presentation you will have a thesis (and that thesis may help to
confirm or contradict a prior hypothesis).

## Types of Argumentation

Within logical arguments, there are two subtypes that
have a place in making arguments from data. These are:

- **deductive reasoning**
- **inductive reasoning**

In deductive reasoning we start with general assumptions and
show that certain conclusions logically follow from them. A
classic example is:

> Socrates is a man. All men are mortal. Therefore, Socrates is
> mortal.

If the assumptions of this statement (first two sentences) are
true, the conclusions **must** be true.

Inductive reasoning, in contrast, builds a conclusion by inferring
based on patterns seen in particular examples. For instance:

> I have taught a total of 700 students over the past 5 years. I
> enjoyed teaching all 700 of them. Therefore, I enjoy teaching all
> students.

While the data provides strong evidence for the conclusion, it does
not guarantee its validity even if the assumptions and logic is
infallible.

## Data-Driven Logical Arguments

Deductive reasoning occurs in statistics when some of our
facts (assumptions) are derived from an analysis of a dataset.
Generally this occurs when we are drawing data from a particular
population. For example, say we are looking at election results
from every county in the United States. The following is a
deductive argument:

> A presidential candidate that has more than 270 electoral votes
> wins the election (assumption). Candidate A had 300 elector votes
> in 2020 (assumption derived from data). Therefore, candidate A
> won the election.

Notice that often not **all** facts are derived from data, but
often some of them are. Inductive reasoning usually occurs in
statistics when sampling from a larger population or observing a
random process. For example:

> Only 1 of the 1000 patients injected with the vaccine had serious
> side-effects. Therefore, the vaccine is safe for distribution.

While inductive reasoning is more traditionally associated with
statistics, the deductive case is quite common in both industry
and academia.

## Examples

If you would like some good examples of data-driven arguments, check out
some of these articles:

- [Hollywood Studios Barely Promoted Non-White Actors And Films](https://fivethirtyeight.com/features/oscars-diversity-hollywood-ad-campaign/)
- [The Minimum Wage Movement is Learning Tipped Workers Behind](https://fivethirtyeight.com/features/the-minimum-wage-movement-is-leaving-tipped-workers-behind/)
- [What Would Happen If We Just Gave People Money?](http://fivethirtyeight.com/features/universal-basic-income/)
- [What Went From in the Flight Water Crisis in Michigan](https://fivethirtyeight.com/features/what-went-wrong-in-flint-water-crisis-michigan/) (Just read Part I)
- [Shut Up About Harvard](https://fivethirtyeight.com/features/shut-up-about-harvard/)
- [How Trump's Supreme Court Could Overturn Roe v. Wade Without Overturning It](http://fivethirtyeight.com/features/how-trumps-supreme-court-could-overturn-roe-v-wade-without-overturning-it/)
- [Tom Brady Doesn't Need Gronk or Moss or Welker to Win](http://fivethirtyeight.com/features/tom-brady-doesnt-need-gronk-or-moss-or-welker-to-win/)
- [A Plagiarism Scandal Is Unfolding In The Crossword World](http://fivethirtyeight.com/features/a-plagiarism-scandal-is-unfolding-in-the-crossword-world/)
- [The Chris Paul Conundrum](https://fivethirtyeight.com/features/the-chris-paul-conundrum/)
- [Under a New System Clinton Could Have Won the Popular Vote by 5 Points and Still Lost](http://fivethirtyeight.com/features/under-a-new-system-clinton-could-have-won-the-popular-vote-by-5-points-and-still-lost/)



