---
title: "Class 17: Exam III Review"
author: "Taylor Arnold"
output: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "class17/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(smodels)
library(ggplot2)
theme_set(theme_minimal())
```

## Format

As with Exam II, this exam will have two parts: an in-class exam and a
take-home component. Your grade will be distributed putting 1/2 of the
weight on the in-class portion and 1/2 on the take-home. Unlike the
second exam, the take-home component will be due *after* the in-class exam.

- In-class, April 2nd (1/2)
- Take home, Due April 4th (1/2)

Instructions for completing the take home exam will be provided in the script.

## Study Guide (In-class)

There are a few things you should be able regarding statistical graphics and
data filtering on the exam by memory:

- construct a basic scatter plot from a dataset
- assign a color to a plot by mapping a variables to the color aesthetic
- set color aesthetic to a fixed value
- filter a dataset according to a numeric variable with one of the
operators `==`, `!=`, `>`, `<`, `>=`, and `<=`.

You should also be able to understand, but do not need to be able to produce
all the details from memory, all of the statistical graphics and filtering
commands that we have used. These include:

- assigning a mix of other aesthetics to a plot
- applying `facet_wrap` to a plot
- using the other aesthetics, including shape, size and alpha
- overriding the `x`, `y`, and `data` values from the defaults in the first line
- filter based on multiple conditions
- filter based on date and time values
- add manual annotations to a plot
- add labels and titles to a plot

I will test these through some combination of (i) showing a code snippet and
asking you questions about what the output should look like, (ii) showing
you code and asking you to fill in a minor mission detail, or (iii) showing
you a plot and asking you about the aesthetics that created the plot.
